---
title: "Ledesma-David-HOMEWORK-02"
author: "David Ledesma"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.
```{r}
x <- 0:100
l = 15
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, 
        space = 0, xlab = "x", ylab = "Pr(X = x)", 
        main = paste0("Probability Mass Function\nlambda = ", l))  
```

###1.a. What is the probability that she will hear more than 8 calls during any given session?
```{r}
x <- 0:100
l <- 15
barplot(ppois(q = x, lambda = l), ylim = 0:1, space = 0, names.arg = x, xlab = "x", 
        ylab = "Pr(X ≤ x)", main = paste0("Cumulative Probability\nlambda = ", 
                                          l))
1- ppois( 8, lambda = l)
```

###1.b. What is the probability that she will hear no calls in a session?
```{r}
dpois(0, lambda = l)
```

###1.c. What is the probability that she will hear exactly 3 calls in a session?
```{r}
dpois(3, lambda = l)
```

###1.d. Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
y <- 0:30
l = 15
probset <- dpois(x = y, lambda = l)
barplot(probset, names.arg = y, 
        space = 0, xlab = "x", ylab = "Pr(X = x)", 
        main = paste0("Probability Mass Function\nlambda = ", l))
```

###1.e. Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).


###1.f. Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?