---
title: "Ledesma-David-HOMEWORK-02"
author: "David Ledesma"
date: "2/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.
```{r}
x <- 0:100
l = 15
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, 
        space = 0, xlab = "x", ylab = "Pr(X = x)", 
        main = paste0("Probability Mass Function\nlambda = ", l))  
```

###1.a. What is the probability that she will hear more than 8 calls during any given session?
```{r}
x <- 0:100
l <- 15
barplot(ppois(q = x, lambda = l), ylim = 0:1, space = 0, names.arg = x, xlab = "x", 
        ylab = "Pr(X ≤ x)", main = paste0("Cumulative Probability\nlambda = ", 
                                          l))
1- ppois( 8, lambda = l)
```

###1.b. What is the probability that she will hear no calls in a session?
```{r}
dpois(0, lambda = l)
```

###1.c. What is the probability that she will hear exactly 3 calls in a session?
```{r}
dpois(3, lambda = l)
```

###1.d. Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
y <- 0:30
l = 15
poisdist <- dpois(x = y, lambda = l)
barplot(poisdist, names.arg = y, 
        space = 0, xlab = "x", ylab = "Pr(X = x)", 
        main = paste0("Probability Mass Function\nlambda = ", l))
```

###1.e. Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
```{r}
simulation <- rpois(104, l)
```

###1.f. Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above? Similar but more noticably right skewed.
```{r}
hist(simulation, xlim = c(0,30))
```

```{r}
y <- 0:30
l = 15
poisdist <- dpois(x = y, lambda = l)
barplot(poisdist, names.arg = y, 
        space = 0, xlab = "x", ylab = "Pr(X = x)", 
        main = paste0("Probability Mass Function\nlambda = ", l))
```
####vs.
probability mass function

#Question 2


```{r}
library(curl)
f <- f <- curl("https://raw.githubusercontent.com/difiore/ADA-2019/master/zombies.csv")
d <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
head(d)
```

### Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education).

```{r}
pop_mean <- function(x) {sum(x)/length(x)}
standdev <- function(x) {
  sqrt(sum((x - mean(x))^2)/length(x))}

hgt<-pop_mean(d$height)
wgt<-pop_mean(d$weight)
ag<-pop_mean(d$age)
zom<-pop_mean(d$zombies_killed)
ed<-pop_mean(d$years_of_education)

hgt2<-standdev(d$height)
wgt2<-standdev(d$weight)
ag2<-standdev(d$age)
zom2<-standdev(d$zombies_killed)
ed2<-standdev(d$years_of_education)
```

###2.a. Use {ggplot} and make boxplots of each of these variable by gender.
```{r}
library(ggplot2)
library(ggpubr)
theme_set(theme_pubr())

ggpot<- d[, c(4, 5, 6, 7, 8, 10)]

?geom_boxplot
par(mfrow = c(2, 3))
phgt<-ggplot(ggpot, aes(x=NULL, height, color=gender)) +
  geom_boxplot()
phgt
pwgt<-ggplot(ggpot, aes(x=NULL, weight, color=gender)) +
  geom_boxplot()
pwgt
pag<-ggplot(ggpot, aes(x=NULL, age, color=gender)) +
  geom_boxplot()
pag
pzom<-ggplot(ggpot, aes(x=NULL, zombies_killed, color=gender)) +
  geom_boxplot()
pzom
ped<-ggplot(ggpot, aes(x=NULL, years_of_education, color=gender)) +
  geom_boxplot()
ped

figure <- ggarrange(phgt, pwgt, pag, pzom, ped,
                    labels = c("Height", "Weight", "Age", "Zombies killed", "Years educated"),
                    ncol = 2, nrow = 3)
figure
```

### 2.c.Use {ggplot} and make scatterplots of height and weight in relation to age. 
```{r}
twoc<-ggplot(d, aes(x=weight, y=height, color=age)) + geom_point()
twoc
```

### Do these variables seem to be related? In what way? Weight and height have a linear relationship. It also appears that on average older individuals are taller and weight more. 

## 2.d.Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. 

```{r}
lhgt<-length(d$height)
rhgt <- rnorm(lhgt, hgt, hgt2)
par(mfrow = c(2, 2))
hist(rhgt, probability = TRUE)
hist(d$height, probability = TRUE)
qqnorm(d$height, main = "Normal QQ plot Height")
qqline(d$height, col = "gray")

lwgt<-length(d$weight)
rwgt <- rnorm(lwgt, wgt, wgt2)
par(mfrow = c(2, 2))
hist(rwgt, probability = TRUE)
hist(d$weight, probability = TRUE)
qqnorm(d$weight, main = "Normal QQ plot Weight")
qqline(d$weight, col = "gray")

lage<-length(d$age)
rag <- rnorm(lage, ag, ag2)
par(mfrow = c(2, 2))
hist(rag, probability = TRUE)
hist(d$age, probability = TRUE)
qqnorm(d$age, main = "Normal QQ plot Age")
qqline(d$age, col = "gray")

lzom<-length(d$zombies_killed)
rzom <- rnorm(lzom, zom, zom2)
par(mfrow = c(2, 2))
hist(rzom, probability = TRUE)
hist(d$zombies_killed, probability = TRUE)
qqnorm(d$zombies_killed, main = "Normal QQ plot Zombies killed")
qqline(d$zombies_killed, col = "gray")

led<-length(d$years_of_education)
red <- rnorm(led, ed, ed2)
par(mfrow = c(2, 2))
hist(red, probability = TRUE)
hist(d$years_of_education, probability = TRUE)
qqnorm(d$years_of_education, main = "Normal QQ plot Years educated")
qqline(d$years_of_education, col = "gray")
```

###Which seem to be and which do not? The numer of zombied killed and the years of education do not seem to be normally distributed.

##2.e. Now use the sample() function to sample ONE subset of 30 zombies (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you will need to base your estimate of the CIs on some different distribution!

```{r}
library(dplyr)

zomsamp<-sample_n(d, 30, replace = FALSE) #sample ONE subset of 30 zombies (without replacement)

zsv <- function(x) {
  sum((x - mean(x))^2)/length(x)-1
  } #variation
zse <- function(x) {
  sqrt(zsv(x))/length(x)
  }#standard error

zsd <- function(x) {
  sqrt(sum((x - mean(x))^2)/length(x)-1)
  } #standard deviation

#Height confidence interval
zhgt<-mean(zomsamp$height)
zhgt2<-zsd(zomsamp$height)
upper <- zhgt + qnorm(0.975, mean = 0, sd = zhgt2) * zse(zomsamp$heigh)
lower <-zhgt + qnorm(0.025, mean = 0, sd = zhgt2) * zse(zomsamp$heigh)
cihgt <- c(lower, upper)
cihgt
#Weight confidence interval
zwgt<-mean(zomsamp$weight)
zwgt2<-zsd(zomsamp$weight)
upper <- zwgt + qnorm(0.975, mean = 0, sd = zwgt2) * zse(zomsamp$weight)
lower <-zwgt + qnorm(0.025, mean = 0, sd = zwgt2) * zse(zomsamp$weight)
ciwgt <- c(lower, upper)
ciwgt
#Age confidence interval
zag<-mean(zomsamp$age)
zag2<-zsd(zomsamp$age)
upper <- zag + qnorm(0.975, mean = 0, sd = zag2) * zse(zomsamp$age)
lower <-zag + qnorm(0.025, mean = 0, sd = zag2) * zse(zomsamp$age)
cizag <- c(lower, upper)
cizag
```

```{r}
#Number of Zombies killed confidence interval
zzom<-mean(zomsamp$zombies_killed)
zzom2<-zsd(zomsamp$zombies_killed)
quantile(zomsamp$zombies_killed, c(0.025, 0.975))
zse(zomsamp$zombies_killed)
sim <- NULL  
n <- 1
for (i in 1:10000) {
  sim[i] <- sample(zomsamp$zombies_killed, n, replace = TRUE)
  } #CI with bootstrapping
normalCI = function(x, CIlevel = 0.95) {
  ci = mean(zomsamp$zombies_killed) + c(-1, 1) * qnorm(1 - (1 - CIlevel)/2) * sqrt(var(zomsamp$zombies_killed)/length(zomsamp$zombies_killed))
  return(ci) 
  } #normal cI function
hist(sim, xlab = "Mean of x", main = "Distribution of Means of Simulated Samples")
abline(v = mean(sim), col = "blue", lwd = 3)  # mean of our simulated samples
abline(v = mean(zomsamp$zombies_killed), col = "red", lwd = 3)  # mean of our actual vector x
abline(v = quantile(sim, 0.025), col = "blue")  # lower ci bound inferred by simulation
abline(v = quantile(sim, 0.975), col = "blue")  # upper ci bound inferred by simulation
text(x = quantile(sim, 0.025) + 0.25, y = 1000, "lower CI based on simulation", 
     col = "blue", srt = 90, pos = 3)
text(x = quantile(sim, 0.975) - 0.25, y = 1000, "upper CI based on simulation", 
     col = "blue", srt = 90, pos = 3)
quantile(sim, c(0.025, 0.975))  # the quantile() function returns the value of x for a specified quantile
ci <- normalCI(zomsamp$zombies_killed, 0.95) 
abline(v = ci[1], col = "red")  # calculated lower ci bound based on the se of our vector and assuming that our observations are drawn from a normal distribution
abline(v = ci[2], col = "red")  # calculated upper ci bound based on the se of our vector and assuming that our observations are drawn from a normal distribution
text(x = ci[1] - 0.25, y = 1000, "lower CI based on normal distribution", col = "red", 
     srt = 90, pos = 3)
text(x = ci[2] + 0.25, y = 1000, "upper CI based on normal distribution", col = "red", 
     srt = 90, pos = 3)
```

```{r}
#Years educated confidence interval
zed<-mean(zomsamp$years_of_education)
zed2<-zsd(zomsamp$years_of_education)
quantile(zomsamp$years_of_education, c(0.025, 0.975))
zse(zomsamp$years_of_education)
sim2 <- NULL  
n <- 1
for (i in 1:10000) {
  sim2[i] <- sample(zomsamp$years_of_education, n, replace = TRUE)
  }#CI with bootstrapping
normalCI = function(x, CIlevel = 0.95) {
  ci = mean(zomsamp$years_of_education) + c(-1, 1) * qnorm(1 - (1 - CIlevel)/2) * sqrt(var(zomsamp$years_of_education)/length(zomsamp$years_of_education))
  return(ci)
  }
hist(sim2, xlab = "Mean of x", main = "Distribution of Means of Simulated Samples")
abline(v = mean(sim2), col = "blue", lwd = 3)  # mean of our simulated samples
abline(v = mean(zomsamp$years_of_education), col = "red", lwd = 3)  # mean of our actual vector x
abline(v = quantile(sim2, 0.025), col = "blue")  # lower ci bound inferred by simulation
abline(v = quantile(sim2, 0.975), col = "blue")  # upper ci bound inferred by simulation
text(x = quantile(sim2, 0.025) + 0.25, y = 1000, "lower CI based on simulation", 
     col = "blue", srt = 90, pos = 3)
text(x = quantile(sim, 0.975) - 0.25, y = 1000, "upper CI based on simulation", 
     col = "blue", srt = 90, pos = 3)
quantile(sim2, c(0.025, 0.975))  # the quantile() function returns the value of x for a specified quantile
ci <- normalCI(zomsamp$years_of_education, 0.95) 
abline(v = ci[1], col = "red")  # calculated lower ci bound based on the se of our vector and assuming that our observations are drawn from a normal distribution
abline(v = ci[2], col = "red")  # calculated upper ci bound based on the se of our vector and assuming that our observations are drawn from a normal distribution
text(x = ci[1] - 0.25, y = 1000, "lower CI based on normal distribution", col = "red", 
     srt = 90, pos = 3)
text(x = ci[2] + 0.25, y = 1000, "upper CI based on normal distribution", col = "red", 
     srt = 90, pos = 3)
```

##2.f. Now draw 99 more random samples of 30 zombies out and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of the sampling distribution for each variable? How do the standard deviations compare to the standard errors estimated in [2.e.]? What do these sampling distributions look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

```{r}
#draw 99 more random samples of 30 zombies out
zomsamp2<-NULL
zomsamp2[[i]]<-NULL
for (i in c(1:99)) {
  zomsamp2[[i]]<- sample_n(d, 30, replace = TRUE)
  }  
```
###calculate the mean for each of the these samples
```{r}

zomsamp2<-as.data.frame(zomsamp2)
wow<-dplyr::select_if(zomsamp2, is.numeric)
wow<-sapply(wow, mean) #takes mean
wow<-t(wow) #transpose
wow<-as.data.frame(wow)
zomweight<- wow[,grep("weight", colnames(wow))] 
zomweight #mean weight
zomheight<-wow[,grep("height", colnames(wow))] 
zomheight #mean height
zomag<-wow[,grep("age", colnames(wow))] 
zomag #mean age
zomzom<-wow[,grep("zombies_killed", colnames(wow))] 
zomzom #mean zombies killed
zomed<-wow[,grep("years_of_education", colnames(wow))] 
zomed#mean years educated
```
###What are the means and standard deviations of the sampling distribution for each variable?
```{r}
#What are the means and standard deviations of the sampling distribution for each variable?
library(tidyverse)
library(dplyr)
library(tidyr)
#totals for each variable, old+new
tzomweight<-sum(zomsamp2[,grep("weight", colnames(zomsamp2))]) + sum(zomsamp$weight)
tzomheight<-sum(zomsamp2[,grep("height", colnames(zomsamp2))]) + sum(zomsamp$height) 
tzomag<-sum(zomsamp2[,grep("age", colnames(zomsamp2))]) + sum(zomsamp$age) 
tzomzom<-sum(zomsamp2[,grep("zombies_killed", colnames(zomsamp2))]) + sum(zomsamp$zombies_killed)
tzomed<-sum(zomsamp2[,grep("years_of_education", colnames(zomsamp2))]) + sum(zomsamp$years_of_education) 

```
### Means for each variable
```{r}

z1<-(tzomweight)/3000
z1 #mean weight
z2<-(tzomheight)/3000
z2 #mean height
z3<-(tzomag)/3000
z3 #mean age
z4<-(tzomzom)/3000
z4 #mean zombies killed
z5<-(tzomed)/3000
z5 #mean years educated
```

###Sd for each variable
```{r}
##unite old+new samples into a matrix
tzomweight2<-cbind(zomsamp2[,grep("weight", colnames(zomsamp2))], zomsamp$weight) 
tzomheight2<-cbind(zomsamp2[,grep("height", colnames(zomsamp2))], zomsamp$height)
tzomage2<-cbind(zomsamp2[,grep("age", colnames(zomsamp2))], zomsamp$age)
tzomzom2<-cbind(zomsamp2[,grep("zombies_killed", colnames(zomsamp2))], zomsamp$zombies_killed)
tzomed2<-cbind(zomsamp2[,grep("years_of_education", colnames(zomsamp2))], zomsamp$years_of_education) 
#gather into one matrix column
tzomweight2<-gather(tzomweight2)
tzomweight2<-dplyr::select_if(tzomweight2, is.numeric)
tzomweight2<-as.matrix(tzomweight2)
          
tzomheight2<-gather(tzomheight2) 
tzomheight2<-dplyr::select_if(tzomheight2, is.numeric)
tzomheight2<-as.matrix(tzomheight2)

tzomage2<-gather(tzomage2) 
tzomage2<- dplyr::select_if(tzomage2, is.numeric)
tzomage2<-as.matrix(tzomage2)

tzomzom2<-gather(tzomzom2) 
tzomzom2<- dplyr::select_if(tzomzom2, is.numeric)
tzomzom2<-as.matrix(tzomzom2)

tzomed2<-gather(tzomed2) 
tzomed2<-  dplyr::select_if(tzomed2, is.numeric)
tzomed2<-as.matrix(tzomed2)

#Sd funtion
zomsd <- function(x) {
  sqrt(sum((x - mean(x))^2)/100-1)
}
#calculate sd for each variable
z6<-zomsd(tzomweight2)
z6 #sd weight
z7<-zomsd(tzomheight2)
z7 #sd height
z8<-zomsd(tzomage2)
z8 #sd age
z9<-zomsd(tzomzom2)
z9 #sd xombies killed
z10<-zomsd(tzomed2)
z10 #sd education
```
##How do the standard deviations compare to the standard errors estimated in [2.e.]?
```{r}
zse(zomsamp$heigh)
z7

zse(zomsamp$weight)
z6

zse(zomsamp$age)
z8

zse(zomsamp$zombies_killed)
z9

zse(zomsamp$years_of_education)
z10
```
##What do these sampling distributions look like?
```{r}
par(mfrow = c(2, 2))
hist(tzomweight2)
qqnorm(tzomweight2, main = "Normal QQ plot Weight")
qqline(tzomweight2, col = "gray")

par(mfrow = c(2, 2))
hist(tzomheight2)
qqnorm(tzomheight2, main = "Normal QQ plot Height")
qqline(tzomheight2, col = "gray")

par(mfrow = c(2, 2))
hist(tzomage2)
qqnorm(tzomage2, main = "Normal QQ plot Age")
qqline(tzomage2, col = "gray")

par(mfrow = c(2, 2))
hist(tzomzom2)
qqnorm(tzomzom2, main = "Normal QQ plot Zombies killed")
qqline(tzomzom2, col = "gray")

par(mfrow = c(2, 2))
hist(tzomed2)
qqnorm(tzomed2, main = "Normal QQ plot Education")
qqline(tzomed2, col = "gray")
```

##Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

##All are normally distributed except for the years educated and the number of zombies killed.




