---
title: "Ledesma-David-ADA-Homework-01"
author: "David Ledesma"
date: "1/31/2019"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Homework 01

# Challenge 1
###First, create a vector of 1 word character strings comprising the first line of the Gettysburg address

“Four score and seven years ago our fathers brought forth on this continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal.”

```{r}
getty <- "Four score and seven years ago our fathers brought forth on this continent, a new nation, conceived in Liberty, and dedicated to the proposition that all men are created equal."

getty <- gsub("[[:punct:]]","", getty)

getty <- strsplit(getty, "[[:space:]]")
```

###Then, extract every third element of this vector.

```{r}
y <- unlist(getty)

y[seq(from = 1, to = 30, by = 3)]
```


# Challenge 2
```{r}
print(m <- matrix(data = 1:80, nrow = 8, ncol = 10, byrow = FALSE))
```

###Given the matrix, m, above, extract the 2nd, 3rd, and 6th columns and assign them to the variable x

```{r}
w <- m[ , 2] 
y<- m[ , 3]
z <- m[ , 6]
print(x <- cbind(w, y, z))
```
###Given the matrix, m, above, extract the 6th to 8th row and assign them to the variable x
```{r}
a <- m[ 6, ] 
b<- m[ 8, ]
print(x <- rbind(a, b))
```
###Given the matrix, m, above, extract the elements from row 2, column 2 to row 6, column 9 and assign them to the variable x
```{r}
c <- m[2, 2]
d <- m[6, 9]
print(x <- rbind(c, d))
```


#Challenge 3

###Construct a 4-dimensional, 400 element array (5 x 5 x 4 x 4) named a consisting of the numbers 400 to 1 (i.e., a descending series)
```{r}
a <- array(data = 400:1, dim = c(5, 5, 4, 4))

print(a[1, 1, 1, 2])
print(a[2, 3, 2, ])
print(a[1:5, 1:5, 3, 3])
```


